---
title: "R Notebook - Integration of FD & WT Whole Samples"
output: 
  html_notebook:
    toc: true
    toc_float: true
---

This pipeline integrates 4 "FD" Fibrous Dysplasia and 4 "WT" control samples into one object using Seurat's integration workflow with the Harmony package.  

See:  
https://satijalab.org/seurat/archive/v4.3/integration_introduction  
https://htmlpreview.github.io/?https://github.com/satijalab/seurat.wrappers/blob/master/docs/harmony.html

All samples were run through QC and hematopoietic lineage cell removal in previous steps and the resulting object is used as the input for this pipeline.  

```{r}
# Load libraries.
library(Seurat)
library(harmony)
library(ggplot2)
library(patchwork)
library(R.filesets)
```

```{r}
# Load dataset object.  Variable name is fdpost.combined.
load(file='./Objects/fdpostcombined.rda')
```

# Setup the Seurat Objects
```{r}
# split the dataset into a list of seurat objects
sample.list <- SplitObject(fdpost.combined, split.by = "orig.ident")

# normalize and identify variable features for each dataset independently
sample.list <- lapply(X = sample.list, FUN = function(x) {
    DefaultAssay(x) <- 'RNA'
    x <- NormalizeData(x)
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 3000)
})

# select features that are repeatedly variable across datasets for integration
features <- SelectIntegrationFeatures(object.list = sample.list, nfeatures = 3000)
```

# Perform Integration
```{r}
sample.anchors <- FindIntegrationAnchors(object.list = sample.list, anchor.features = features)
```

```{r}
# this command creates an 'integrated' data assay
fdpost.combined <- IntegrateData(anchorset = sample.anchors)
```
# Perform an Integrated Analysis
```{r}
# specify that we will perform downstream analysis on the corrected data note that the
# original unmodified data still resides in the 'RNA' assay
DefaultAssay(fdpost.combined) <- "integrated"

# Run the standard workflow for visualization and clustering
fdpost.combined <- ScaleData(fdpost.combined, verbose = FALSE, vars.to.regress = 'percent.mt')
fdpost.combined <- RunPCA(fdpost.combined, npcs = 30, verbose = FALSE)

# Commented steps below not needed prior to running Harmony
#fdpost.combined <- RunUMAP(fdpost.combined, reduction = "pca", dims = 1:30, verbose = F)
#fdpost.combined <- FindNeighbors(fdpost.combined, reduction = "pca", dims = 1:30)
#fdpost.combined <- FindClusters(fdpost.combined, resolution = 0.4)
```

# Run Harmony
```{r}
fdpost.combined <- RunHarmony(fdpost.combined, group.by.vars = "orig.ident", 
                              assay.use = 'integrated',
                              reduction = 'pca')
```

## 10 Harmony Dims, Res: 0.4  
```{r}
# 10 Harmony Dims, Res: 0.4.
DefaultAssay(fdpost.combined) <- 'integrated'
fdpost.combined <- FindNeighbors(fdpost.combined, dims = 1:10, verbose=F, 
                                 reduction = 'harmony')
fdpost.combined <- FindClusters(fdpost.combined, resolution = 0.4, verbose=F)
fdpost.combined <- RunUMAP(fdpost.combined, reduction = "harmony", dims = 1:10, verbose=F)
```

## UMAP  
```{r}
DimPlot(fdpost.combined, label = T) + ggtitle('Dims:10, Res:0.4')
```

## Markers: ACTA2, COL1A1, PECAM1, CENPF  
```{r}
# Marker genes for each cluster group.
# ACTA2: Smooth muscle
# COL1A1: Stromal
# PECAM1: Endothelial
# CENPF: Cycling
DefaultAssay(fdpost.combined) <- 'RNA'
FeaturePlot(fdpost.combined, features = c("ACTA2", "PECAM1", "COL1A1", "CENPF"), 
            order = T, min.cutoff = c('q10','q10','q10','q10'))
```

# Save Object
```{r}
saveRDS(fdpost.combined, file='./Objects/fdpostcombined.rds')
```

