---
title: "R Notebook - Cluster Stromal Portions of FD and WT samples using Liger"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

This pipeline clusters the stromal portions of the dataset samples using Seurat and Liger R packages.  The stromal clusters are first subset from the integrated object generated in a previous pipleine and then re-integrated and clustered.

See:  
https://htmlpreview.github.io/?https://github.com/satijalab/seurat.wrappers/blob/master/docs/liger.html  

```{r}
library(rliger)
library(Seurat)
library(SeuratWrappers)
library(R.filesets)
```

# Load Object
```{r}
fdpost.combined <- readRDS(file='./Objects/fdpostcombined.rds')
```

```{r}
DimPlot(fdpost.combined, label = T) + ggtitle('fdpost.combined')
```

# Subset Clusters
Subset stromal clusters (including stromal cycling cells) into their own object.  
Note: When running the next line of code, make sure the cluster numbers given in the idents parameter match the stromal cluster numbers in your own object.  
```{r}
strom.liger <- subset(fdpost.combined, idents = c(0:1,6:10))
```

```{r}
DefaultAssay(strom.liger)
```

```{r}
rm(fdpost.combined)
```

# Normalize and Scale Data
```{r}
strom.liger <- NormalizeData(strom.liger)
strom.liger <- FindVariableFeatures(strom.liger)
```

```{r}
strom.liger <- ScaleData(strom.liger, split.by = "orig.ident", do.center = FALSE)
```

# Run Liger Wrapper Functions
```{r}
strom.liger <- RunOptimizeALS(strom.liger, k = 20, lambda = 4, split.by = "orig.ident")
```
```{r}
strom.liger <- RunQuantileNorm(strom.liger, split.by = "orig.ident")
```

# Cluster Cells
```{r}
# Perform Louvain clustering (`FindNeighbors` and `FindClusters`) after
# `RunQuantileNorm`
strom.liger <- FindNeighbors(strom.liger, reduction = "iNMF", dims = 1:20)
strom.liger <- FindClusters(strom.liger, resolution = 0.45)
# Dimensional reduction and plotting
strom.liger <- RunUMAP(strom.liger, dims = 1:ncol(strom.liger[["iNMF"]]), reduction = "iNMF")
```

## UMAP
```{r}
DimPlot(strom.liger, label = T)
```

# Save Object
```{r}
saveRDS(strom.liger, file="./Objects/stromliger.rds")
```



