---
title: "R Notebook - SoupX on Bone FD samples."
---

This pipeline removes ambient or background mRNA contamination (the "Soup") from single cell RNA-seq samples using the R package, [SoupX](https://www.biorxiv.org/content/10.1101/303727v2).

```{r}
library(SoupX)
library(ggplot2)
```

# Load raw counts  
Load raw counts and clustering data generated from 10X Cell Ranger pipeline into SoupX objects
```{r}
spx.fd.list <- list(
  FD16ps = load10X('../cellranger/FD/FD16-POST/'),
  FD21ps = load10X('../cellranger/FD/FD21-POST/'),
  FD28ps = load10X('../cellranger/FD/FD28-POST/'),
  FD716ps = load10X('../cellranger/FD/FD716-POST/')
)
```

# Probability density distribution plots of contamination fraction
```{r}
for (i in length(spx.fd.list)) {
  print(spx.fd.list[i])
  spx.fd.list[[i]] <- autoEstCont(spx.fd.list[[i]])
}
```

# Tables of top20 contaminating gene reads
```{r}
for (i in length(spx.fd.list)) {
  head(spx.fd.list[[i]]$soupProfile[order(spx.fd.list[[i]]$soupProfile$est, decreasing = TRUE), ], 
       n = 20)
}
```

# Marker distribution plots
```{r}
for (i in length(spx.fd.list)) {
  plotMarkerDistribution(spx.fd.list[[i]]) + ggtitle(names(spx.fd.list[i]))
}
```

# Adjust counts
```{r}
spx.out <- list()
for (i in length(spx.fd.list)) {
  spx.fd.out[[names(spx.fd.list[i])]] = adjustCounts(spx.fd.list[[i]], roundToInt = TRUE)
}
```

# Save output
```{r}
save(spx.fd.out, file='./Objects/SoupX_FD.rda')
```



