---
title: "R Notebook: Remove High Mito% Cells from FD Samples"
---

This script visualizes the mitochondrial read percentages in each sample and removes cells with greater than or equal to 15% of reads coming from mitochondrial genes.

```{r}
library(Seurat)
library(ggplot2)
library(patchwork)
```

# Load adjusted counts from SoupX pipeline.
```{r}
# loads spx.fd.out object
load(file='./Objects/SoupX_FD.rda')
```

# Create Seurat Objects
```{r}
FD16ps <- CreateSeuratObject(counts = spx.fd.out[["FD16ps"]], project = "FD16ps", 
                             min.cells = 3, min.features = 200)
FD21ps <- CreateSeuratObject(counts = spx.fd.out[["FD21ps"]], project = "FD21ps", 
                             min.cells = 3, min.features = 200)
FD28ps <- CreateSeuratObject(counts = spx.fd.out[["FD28ps"]], project = "FD28ps", 
                             min.cells = 3, min.features = 200)
FD716ps <- CreateSeuratObject(counts = spx.fd.out[["FD716ps"]], project = "FD716ps", 
                              min.cells = 3, min.features = 200)
```

```{r}
rm(spx.fd.out)
```


# Calculate mitochondrial percentage and store in percent.mt variable
```{r}
FD16ps[["percent.mt"]] <- PercentageFeatureSet(object = FD16ps, pattern = "^MT-")
FD21ps[["percent.mt"]] <- PercentageFeatureSet(object = FD21ps, pattern = "^MT-")
FD28ps[["percent.mt"]] <- PercentageFeatureSet(object = FD28ps, pattern = "^MT-")
FD716ps[["percent.mt"]] <- PercentageFeatureSet(object = FD716ps, pattern = "^MT-")
```

# Violin Plots of percent.mt
## Linear Scale
```{r}
(VlnPlot(FD16ps, features = "percent.mt") + 
  VlnPlot(FD21ps, features = "percent.mt") +
  VlnPlot(FD28ps, features = "percent.mt") + 
  VlnPlot(FD716ps, features = "percent.mt")) & NoLegend()
```

## Log Scale
```{r}
(VlnPlot(FD16ps, features = "percent.mt", log=TRUE) + 
  VlnPlot(FD21ps, features = "percent.mt", log=TRUE) +
  VlnPlot(FD28ps, features = "percent.mt", log=TRUE) + 
  VlnPlot(FD716ps, features = "percent.mt", log=TRUE)) & NoLegend()
```

## y-axis range: 0-25
Also, display max density or mito% value where the violin plot is the widest (i.e. has the greatest number of cells).
```{r}
( (VlnPlot(FD16ps, features = "percent.mt") + 
     labs(subtitle = paste("Max Density: ",
                           toString(round(density(FD16ps$percent.mt)$x[which.max(
                             density(FD16ps$percent.mt)$y)], digits = 2)))) ) + 
    (VlnPlot(FD21ps, features = "percent.mt") +
       labs(subtitle = paste("Max Density: ",
                             toString(round(density(FD21ps$percent.mt)$x[which.max(
                               density(FD21ps$percent.mt)$y)], digits = 2)))) ) + 
    
    (VlnPlot(FD28ps, features = "percent.mt") +
       labs(subtitle = paste("Max Density: ",
                             toString(round(density(FD28ps$percent.mt)$x[which.max(
                               density(FD28ps$percent.mt)$y)], digits = 2)))) ) + 
    
    (VlnPlot(FD716ps, features = "percent.mt") +
       labs(subtitle = paste("Max Density: ",
                             toString(round(density(FD716ps$percent.mt)$x[which.max(
                               density(FD716ps$percent.mt)$y)], digits = 2)))) ) ) & 
  NoLegend() & ylim(0,25)
```

# Show number of cells that will be removed per sample
```{r}
table(FD16ps$percent.mt > 15)
table(FD21ps$percent.mt > 15)
table(FD28ps$percent.mt > 15)
table(FD716ps$percent.mt > 15)
```

# Keep cells with < 15 percent.mt
```{r}
FD16ps <- subset(FD16ps, subset = percent.mt < 15)
FD21ps <- subset(FD21ps, subset = percent.mt < 15)
FD28ps <- subset(FD28ps, subset = percent.mt < 15)
FD716ps <- subset(FD716ps, subset = percent.mt < 15)
```

# Violin Plots after filtering
```{r}
(VlnPlot(FD16ps, features = "percent.mt") + 
  VlnPlot(FD21ps, features = "percent.mt") +
  VlnPlot(FD28ps, features = "percent.mt") + 
  VlnPlot(FD716ps, features = "percent.mt")) & NoLegend()
```

# Save objects
```{r}
save(FD16ps, FD21ps, FD28ps, FD716ps, file='./Objects/postMT_FD.rda')
```