---
title: "R Notebook: Remove High Mito% Cells from WT Samples"
---

This script visualizes the mitochondrial read percentages in each sample and removes cells with greater than or equal to 15% of reads coming from mitochondrial genes.

```{r}
library(Seurat)
library(ggplot2)
library(patchwork)
```

# Load adjusted counts from SoupX pipeline.
```{r}
# loads spx.wt.out object
load(file='./Objects/SoupX_WT.rda')
```

# Create Seurat Objects
```{r}
WT150ps <- CreateSeuratObject(counts = spx.wt.out[["WT150ps"]], project = "WT150ps", 
                             min.cells = 3, min.features = 200)
WT151ps <- CreateSeuratObject(counts = spx.wt.out[["WT151ps"]], project = "WT151ps", 
                             min.cells = 3, min.features = 200)
WT152ps <- CreateSeuratObject(counts = spx.wt.out[["WT152ps"]], project = "WT152ps", 
                             min.cells = 3, min.features = 200)
WT153ps <- CreateSeuratObject(counts = spx.wt.out[["WT153ps"]], project = "WT153ps", 
                              min.cells = 3, min.features = 200)
```

```{r}
rm(spx.wt.out)
```


# Calculate mitochondrial percentage and store in percent.mt variable
```{r}
WT150ps[["percent.mt"]] <- PercentageFeatureSet(object = WT150ps, pattern = "^MT-")
WT151ps[["percent.mt"]] <- PercentageFeatureSet(object = WT151ps, pattern = "^MT-")
WT152ps[["percent.mt"]] <- PercentageFeatureSet(object = WT152ps, pattern = "^MT-")
WT153ps[["percent.mt"]] <- PercentageFeatureSet(object = WT153ps, pattern = "^MT-")
```

# Violin Plots of percent.mt
## Linear Scale
```{r}
(VlnPlot(WT150ps, features = "percent.mt") + 
  VlnPlot(WT151ps, features = "percent.mt") +
  VlnPlot(WT152ps, features = "percent.mt") + 
  VlnPlot(WT153ps, features = "percent.mt")) & NoLegend()
```

## Log Scale
```{r}
(VlnPlot(WT150ps, features = "percent.mt", log=TRUE) + 
  VlnPlot(WT151ps, features = "percent.mt", log=TRUE) +
  VlnPlot(WT152ps, features = "percent.mt", log=TRUE) + 
  VlnPlot(WT153ps, features = "percent.mt", log=TRUE)) & NoLegend()
```

## y-axis range: 0-25
Also, display max density or mito% value where the violin plot is the widest (i.e. has the greatest number of cells).
```{r}
( (VlnPlot(WT150ps, features = "percent.mt") + 
     labs(subtitle = paste("Max Density: ",
                           toString(round(density(WT150ps$percent.mt)$x[which.max(
                             density(WT150ps$percent.mt)$y)], digits = 2)))) ) + 
    (VlnPlot(WT151ps, features = "percent.mt") +
       labs(subtitle = paste("Max Density: ",
                             toString(round(density(WT151ps$percent.mt)$x[which.max(
                               density(WT151ps$percent.mt)$y)], digits = 2)))) ) + 
    
    (VlnPlot(WT152ps, features = "percent.mt") +
       labs(subtitle = paste("Max Density: ",
                             toString(round(density(WT152ps$percent.mt)$x[which.max(
                               density(WT152ps$percent.mt)$y)], digits = 2)))) ) + 
    
    (VlnPlot(WT153ps, features = "percent.mt") +
       labs(subtitle = paste("Max Density: ",
                             toString(round(density(WT153ps$percent.mt)$x[which.max(
                               density(WT153ps$percent.mt)$y)], digits = 2)))) ) ) & 
  NoLegend() & ylim(0,25)
```

# Show number of cells that will be removed per sample
```{r}
table(WT150ps$percent.mt > 15)
table(WT151ps$percent.mt > 15)
table(WT152ps$percent.mt > 15)
table(WT153ps$percent.mt > 15)
```

# Keep cells with < 15 percent.mt
```{r}
WT150ps <- subset(WT150ps, subset = percent.mt < 15)
WT151ps <- subset(WT151ps, subset = percent.mt < 15)
WT152ps <- subset(WT152ps, subset = percent.mt < 15)
WT153ps <- subset(WT153ps, subset = percent.mt < 15)
```

# Violin Plots after filtering
```{r}
(VlnPlot(WT150ps, features = "percent.mt") + 
  VlnPlot(WT151ps, features = "percent.mt") +
  VlnPlot(WT152ps, features = "percent.mt") + 
  VlnPlot(WT153ps, features = "percent.mt")) & NoLegend()
```

# Save objects
```{r}
save(WT150ps, WT151ps, WT152ps, WT153ps, file='./Objects/postMT_WT.rda')
```