---
title: "R Notebook - Remove Gene Count Outlier Cells From FD Samples"
---

This script visualizes cell gene (nFeature_RNA variable) and read (nCount_RNA variable) counts in each sample and removes outlier cells with fewer than 300 or greater than 5500 genes.

```{r}
library(Seurat)
library(ggplot2)
library(patchwork)
```

# Load Seurat objects  
Loads FD sample objects FD16ps, FD21ps, FD28ps, and FD716ps.
```{r}
# Load Seurat objects that have had cells with >= 15% mitochondrial gene reads removed.
load(file='./Objects/postMT_FD.rda')
```

# Violin plots before filtering
```{r}
# Violin plots of nCount_RNA.  Log scale = False.
for(i in c(FD16ps,FD21ps, FD28ps, FD716ps)){
  print(VlnPlot(i, features = "nCount_RNA") + NoLegend())
}
```

```{r}
# Violin plots of nCount_RNA.  Log scale = True.
for(i in c(FD16ps,FD21ps, FD28ps, FD716ps)){
  print(VlnPlot(i, features = "nCount_RNA", log = T) + NoLegend())
}
```

```{r}
# Violin plots of nFeature_RNA.  Log scale = False.
for(i in c(FD16ps,FD21ps, FD28ps, FD716ps)){
  print(VlnPlot(i, features = "nFeature_RNA") + NoLegend())
}
```

```{r}
# Violin plots of nFeature_RNA.  Log scale = True.
for(i in c(FD16ps,FD21ps, FD28ps, FD716ps)){
  print(VlnPlot(i, features = "nFeature_RNA", log = T) + NoLegend())
}
```

# Scatter plots before filtering  
```{r}
# Scatter plots of nCount_RNA vs nFeature_RNA.
for(i in c(FD16ps,FD21ps, FD28ps, FD716ps)){
  print(FeatureScatter(i, feature1 = "nCount_RNA", feature2 = "nFeature_RNA"))
}
```

# Number of cells before filtering
```{r}
c(FD16ps, FD21ps, FD28ps, FD716ps)
```


# Filter cells
```{r}
# Keep cells with > 300 nFeature_RNA & < 8000 nFeature_RNA.
FD16ps <- subset(FD16ps, subset = nFeature_RNA > 300 & nFeature_RNA < 5500)
FD21ps <- subset(FD21ps, subset = nFeature_RNA > 300 & nFeature_RNA < 5500)
FD28ps <- subset(FD28ps, subset = nFeature_RNA > 300 & nFeature_RNA < 5500)
FD716ps <- subset(FD716ps, subset = nFeature_RNA > 300 & nFeature_RNA < 5500)
```

# Number of cells after filtering
```{r}
c(FD16ps, FD21ps, FD28ps, FD716ps)
```

# Violin plots after filtering
```{r}
# Violin plots of nCount_RNA.  Log scale = False.
for(i in c(FD16ps,FD21ps, FD28ps, FD716ps)){
  print(VlnPlot(i, features = "nCount_RNA") + NoLegend())
}
```

```{r}
# Violin plots of nCount_RNA.  Log scale = True.
for(i in c(FD16ps,FD21ps, FD28ps, FD716ps)){
  print(VlnPlot(i, features = "nCount_RNA", log = T) + NoLegend())
}
```

```{r}
# Violin plots of nFeature_RNA.  Log scale = False.
for(i in c(FD16ps,FD21ps, FD28ps, FD716ps)){
  print(VlnPlot(i, features = "nFeature_RNA") + NoLegend())
}
```

```{r}
# Violin plots of nFeature_RNA.  Log scale = True.
for(i in c(FD16ps,FD21ps, FD28ps, FD716ps)){
  print(VlnPlot(i, features = "nFeature_RNA", log = T) + NoLegend())
}
```

# Scatter plots after filtering  
```{r}
# Scatter plots of nCount_RNA vs nFeature_RNA.
for(i in c(FD16ps,FD21ps, FD28ps, FD716ps)){
  print(FeatureScatter(i, feature1 = "nCount_RNA", feature2 = "nFeature_RNA"))
}
```

# Save objects
```{r}
save(FD16ps, FD21ps, FD28ps, FD716ps, file='./Objects/postnFeat_FD.rda')
```