---
title: "R Notebook - Remove Gene Count Outlier Cells From WT Samples"
---

This script visualizes cell gene (nFeature_RNA variable) and read (nCount_RNA variable) counts in each sample and removes outlier cells with fewer than 300 or greater than 5500 genes.

```{r}
library(Seurat)
library(ggplot2)
library(patchwork)
```

# Load Seurat objects  
Loads FD sample objects WT150ps, WT151ps, WT152ps, and WT153ps.
```{r}
# Load Seurat objects that have had cells with >= 15% mitochondrial gene reads removed.
load(file='./Objects/postMT_WT.rda')
```

# Violin plots before filtering
```{r}
# Violin plots of nCount_RNA.  Log scale = False.
for(i in c(WT150ps,WT151ps, WT152ps, WT153ps)){
  print(VlnPlot(i, features = "nCount_RNA") + NoLegend())
}
```

```{r}
# Violin plots of nCount_RNA.  Log scale = True.
for(i in c(WT150ps,WT151ps, WT152ps, WT153ps)){
  print(VlnPlot(i, features = "nCount_RNA", log = T) + NoLegend())
}
```

```{r}
# Violin plots of nFeature_RNA.  Log scale = False.
for(i in c(WT150ps,WT151ps, WT152ps, WT153ps)){
  print(VlnPlot(i, features = "nFeature_RNA") + NoLegend())
}
```

```{r}
# Violin plots of nFeature_RNA.  Log scale = True.
for(i in c(WT150ps,WT151ps, WT152ps, WT153ps)){
  print(VlnPlot(i, features = "nFeature_RNA", log = T) + NoLegend())
}
```

# Scatter plots before filtering  
```{r}
# Scatter plots of nCount_RNA vs nFeature_RNA.
for(i in c(WT150ps,WT151ps, WT152ps, WT153ps)){
  print(FeatureScatter(i, feature1 = "nCount_RNA", feature2 = "nFeature_RNA"))
}
```

# Number of cells before filtering
```{r}
c(WT150ps, WT151ps, WT152ps, WT153ps)
```


# Filter cells
```{r}
# Keep cells with > 300 nFeature_RNA & < 8000 nFeature_RNA.
WT150ps <- subset(WT150ps, subset = nFeature_RNA > 300 & nFeature_RNA < 5500)
WT151ps <- subset(WT151ps, subset = nFeature_RNA > 300 & nFeature_RNA < 5500)
WT152ps <- subset(WT152ps, subset = nFeature_RNA > 300 & nFeature_RNA < 5500)
WT153ps <- subset(WT153ps, subset = nFeature_RNA > 300 & nFeature_RNA < 5500)
```

# Number of cells after filtering
```{r}
c(WT150ps, WT151ps, WT152ps, WT153ps)
```

# Violin plots after filtering
```{r}
# Violin plots of nCount_RNA.  Log scale = False.
for(i in c(WT150ps,WT151ps, WT152ps, WT153ps)){
  print(VlnPlot(i, features = "nCount_RNA") + NoLegend())
}
```

```{r}
# Violin plots of nCount_RNA.  Log scale = True.
for(i in c(WT150ps,WT151ps, WT152ps, WT153ps)){
  print(VlnPlot(i, features = "nCount_RNA", log = T) + NoLegend())
}
```

```{r}
# Violin plots of nFeature_RNA.  Log scale = False.
for(i in c(WT150ps,WT151ps, WT152ps, WT153ps)){
  print(VlnPlot(i, features = "nFeature_RNA") + NoLegend())
}
```

```{r}
# Violin plots of nFeature_RNA.  Log scale = True.
for(i in c(WT150ps,WT151ps, WT152ps, WT153ps)){
  print(VlnPlot(i, features = "nFeature_RNA", log = T) + NoLegend())
}
```

# Scatter plots after filtering  
```{r}
# Scatter plots of nCount_RNA vs nFeature_RNA.
for(i in c(WT150ps,WT151ps, WT152ps, WT153ps)){
  print(FeatureScatter(i, feature1 = "nCount_RNA", feature2 = "nFeature_RNA"))
}
```

# Save objects
```{r}
save(WT150ps, WT151ps, WT152ps, WT153ps, file='./Objects/postnFeat_FD.rda')
```