---
title: "R Notebook - Filtering of Hematopoietic Lineage Cells from Dataset"
output:
  html_notebook:
    toc: true
    toc_float: true
---

The purpose of this pipeline is to remove hematopoietic lineage cells from samples, including peripheral blood mononuclear cells (PBMCs) and red blood cells (RBCs).  

Samples are integrated into a single object, and then the R packages SingleR and Seurat are used to perform reference-based annotation of single cells.  See:  
https://satijalab.org/seurat/archive/v4.3/integration_introduction  
https://github.com/dviraran/singler  
https://satijalab.org/seurat/archive/v4.3/multimodal_reference_mapping

These annotations and the PTPRC (CD45 protein) hematopoietic marker gene are used to filter out hematopoietic lineage cells from the dataset.  Any remaining RBC-contaminated cells are removed by filtering out those with >3 HBB, HBA1, or HBA2 transcripts.  Post-filtering, samples are re-integrated  and cells reclustered in a subsequent pipeline.

The steps in this pipeline are as follows:  
1) Integrate samples into one Seurat object.  
2) Run SingleR's reference mapping pipeline.  
3) Run Seurat's reference mapping pipeline.  
4) Remove blood cells based on SingleR's reference-based annotations.  
5) Remove PTPRC+ (a hemapoietic lineage marker) cells.  
6) Remove blood cells based on Seurat's reference-based annotations.  
7) Remove remaining RBCs and RBC-contanminated cells.  

(Note: For simplicitie's sake, FD28ps is included in the script below.  However, for the published analysis, FD28ps was collected at a later date, processed individually using the same filtering steps and cutoffs before being integrated with the other samples post-filtering.)  

```{r}
# Load libraries
library(Seurat)
library(SeuratDisk)
library(SingleR)
library(ggplot2)
library(patchwork)
```

```{r}
# Load samples
load("./Objects/postnFeat_WT.rda")
load("./Objects/postnFeat_FD.rda")
```

# 1) Integrate Samples  
```{r}
# Normalize data with SCTransform
list.post <- lapply(X = list(FD16ps, FD21ps, FD28ps, FD716ps, 
                             WT150ps, WT151ps, WT152ps, WT153ps), 
                    FUN = SCTransform, 
                    vars.to.regress = c('nCount_RNA','percent.mt'))
```

```{r}
rm(FD16ps, FD21ps, FD28ps, FD716ps, WT150ps, WT151ps, WT152ps, WT153ps)
```

```{r}
features.post <- SelectIntegrationFeatures(object.list = list.post, nfeatures = 3000)
```

```{r}
list.post <- PrepSCTIntegration(object.list = list.post, anchor.features = features.post)
```

```{r}
post.anchors <- FindIntegrationAnchors(object.list = list.post, normalization.method = "SCT", 
    anchor.features = features.post)
```

```{r}
rm(list.post)
```

```{r}
fdpost.combined <- IntegrateData(anchorset = post.anchors, normalization.method = "SCT")
```

```{r}
rm(post.anchors)
```

```{r}
fdpost.combined <- RunPCA(fdpost.combined, npcs = 50)
```

```{r}
fdpost.combined <- RunUMAP(fdpost.combined, dims=1:30)
```

```{r}
fdpost.combined <- FindNeighbors(fdpost.combined, reduction = "pca", dims = 1:30)
```

```{r}
fdpost.combined <- FindClusters(fdpost.combined, resolution = 0.5)
```

```{r}
fdpost.combined
```

```{r}
# Total number of cells in the object
ncol(fdpost.combined)
# Number of cells per sample
table(fdpost.combined$orig.ident)
```

# 2) Run SingleR Pipeline
```{r}
# Load Human Primary Cell Atlas reference  
hpca.se <- celldex::HumanPrimaryCellAtlasData()
```
```{r}
# Load Blueprint Encode reference  
encode.se <- celldex::BlueprintEncodeData()
```

```{r}
# Run SingleR using hpca.se ref.  
snglr.hpca.main <- SingleR(test = as.matrix(GetAssayData(fdpost.combined[["SCT"]], slot = "data")), 
                           ref = hpca.se,
                           labels = hpca.se$label.main)
snglr.hpca.fine <- SingleR(test = as.matrix(GetAssayData(fdpost.combined[["SCT"]], slot = "data")), 
                           ref = hpca.se,
                           labels = hpca.se$label.fine)
```

```{r}  
# Runc SingleR using encode.se ref.
snglr.enc.main <- SingleR(test = as.matrix(GetAssayData(fdpost.combined[["SCT"]], slot = "data")), 
                          ref = encode.se,
                          labels = encode.se$label.main)
snglr.enc.fine <- SingleR(test = as.matrix(GetAssayData(fdpost.combined[["SCT"]], slot = "data")), 
                          ref = encode.se,
                          labels = encode.se$label.fine)
```

```{r}
# Transfer pruned.labels to object
fdpost.combined[["hpca.main.prnd"]] <- snglr.hpca.main$pruned.labels
fdpost.combined[["encode.main.prnd"]] <- snglr.enc.main$pruned.labels
fdpost.combined[["hpca.fine.prnd"]] <- snglr.hpca.fine$pruned.labels
fdpost.combined[["encode.fine.prnd"]] <- snglr.enc.fine$pruned.labels
```

```{r}
# Transfer labels to object
fdpost.combined[["hpca.main"]] <- snglr.hpca.main$labels
fdpost.combined[["encode.main"]] <- snglr.enc.main$labels
fdpost.combined[["hpca.fine"]] <- snglr.hpca.fine$labels
fdpost.combined[["encode.fine"]] <- snglr.enc.fine$labels
```

## Cell Counts from hpca.main
```{r}
table(fdpost.combined$hpca.main)
```
## Cell Counts from encode.main
```{r}
table(fdpost.combined$encode.main)
```

## Cell Counts from hpca.fine
```{r}
table(fdpost.combined$hpca.fine)
```
## Cell Counts from encode.fine
```{r}
table(fdpost.combined$encode.fine)
```

# 3) Run Seurat's Mapping Pipeline 
Identify PBMC's using Seurat's Reference Mapping Pipeline.  
https://satijalab.org/seurat/archive/v4.3/multimodal_reference_mapping   

## Load Reference  
This reference takes up 9.8 GB of space.
```{r}
reference <- LoadH5Seurat("../data/pbmc_multimodal.h5seurat")
```
## Find Transfer Anchors
```{r}
ref.anchors = FindTransferAnchors(
  reference = reference,
  query = fdpost.combined,
  normalization.method = "SCT",
  reference.reduction = "spca",
  dims = 1:50 #Default value is 1:30. Vignette uses 1:50 dims. -TM
) 
```
## Map Query
```{r}
fdpost.combined = MapQuery(
  anchorset = ref.anchors,
  query = fdpost.combined,
  reference = reference,
  refdata = list(
    celltype.l1 = "celltype.l1",
    celltype.l2 = "celltype.l2",
    predicted_ADT = "ADT"
  ),
  reference.reduction = "spca", 
  reduction.model = "wnn.umap"
)
```

## UMAP Plots Using Reference UMAP
UMAP plots showing l1 and l2 predicted celltypes. The cells are placed onto the reference's UMAP.
```{r}
p1 <- DimPlot(fdpost.combined, reduction = "ref.umap", group.by = "predicted.celltype.l1", 
              label = TRUE, label.size = 3, repel = TRUE) #+ NoLegend()
p2 = DimPlot(fdpost.combined, reduction = "ref.umap", group.by = "predicted.celltype.l2", 
             label = TRUE, label.size = 3 ,repel = TRUE) #+ NoLegend()
p1
p2
(p1 + p2) & NoLegend()
```

## Cell Counts from l1 and l2 Predictions
```{r}
table(fdpost.combined$predicted.celltype.l1)
table(fdpost.combined$predicted.celltype.l2)
```


# 4) Filter using SingleR Refs  
```{r}
# Total number of cells in the object
ncol(fdpost.combined)
```

## Filter Out hpca.main Blood Cells  
```{r}
fdpost.combined <-
  subset(
    fdpost.combined,
    subset = hpca.main != "B_cell" & hpca.main != "BM" &
      hpca.main != "BM & Prog." &
      hpca.main != "CMP" & hpca.main != "DC" &
      hpca.main != "Erythroblast" &
      hpca.main != "GMP" &
      hpca.main != "HSC_CD34+" &
      hpca.main != "Macrophage" &
      hpca.main != "MEP" & hpca.main != "Monocyte" &
      hpca.main != "Neutrophils" &
      hpca.main != "NK_cell" &
      hpca.main != "Pre-B_cell_CD34-" &
      hpca.main != "Pro-B_cell_CD34+" &
      hpca.main != "Pro-Myelocyte" &
      hpca.main != "T_cells"
  )
# Total number of cells in the object
ncol(fdpost.combined)
```

## Filter Out encode.main Blood Cells  
```{r}
fdpost.combined <-
  subset(
    fdpost.combined,
    subset = encode.main != "B-cells" & encode.main != "CD4+ T-cells" &
      encode.main != "CD8+ T-cells" &
      encode.main != "Erythrocytes" &
      encode.main != "HSC" &
      encode.main != "Monocytes" & encode.main != "NK cells"
  )
# Total number of cells in the object
ncol(fdpost.combined)
```

## Filter out hpca & encode "fine" Blood Cells  
```{r}
fdpost.combined <-
  subset(
    fdpost.combined,
    subset = hpca.fine != "HSC_CD34+" & hpca.fine != "NK_cell" &
      hpca.fine != "Pro-B_cell_CD34+" &
      hpca.fine != "B_cell:Plasma_cell" &
      hpca.fine != "DC:monocyte-derived:AEC-conditioned" &
      encode.fine != "CD4+ T-cells" &
      encode.fine != "CLP" & encode.fine != "Erythrocytes" &
      encode.fine != "HSC" &
      encode.fine != "Plasma cells" &
      encode.fine != "naive B-cells"
  )
# Total number of cells in the object
ncol(fdpost.combined)
```

## Cell Counts from hpca.main after Filtering
```{r}
table(fdpost.combined$hpca.main)
```
## Cell Counts from encode.main after Filtering
```{r}
table(fdpost.combined$encode.main)
```


# 5) Filter PTPRC+ (CD45+) Cells
```{r}
VlnPlot(fdpost.combined, features = "PTPRC", group.by = "predicted.celltype.l1", 
        slot = "counts") + NoLegend()
```

```{r}
# Remove PTPRC+ (CD45+) cells.
fdpost.combined <- subset(fdpost.combined, subset = PTPRC < 1)
```

```{r}
# Total number of cells in the object
ncol(fdpost.combined)
```


# 6) Filter using Seurat's Ref.  
```{r}
# Filter out l1 blood cells with a score >= 0.8.
# A large number of predicted monocytes, CD4 T cells and dendritic cells have a low
# prediction score, so only remove those with a score greater than or equal to 0.8.  
ncol(fdpost.combined)
fdpost.combined <- subset(fdpost.combined, subset = (predicted.celltype.l1.score >= 0.8 & 
                   predicted.celltype.l1 != 'other'), invert = TRUE)
ncol(fdpost.combined)
```

```{r}
# Filter out low frequency l1 blood cells
fdpost.combined <- subset(fdpost.combined, subset = predicted.celltype.l1 != "CD8 T" & 
                   predicted.celltype.l1 != "other T")
```

```{r}
# Filter out l2 blood cells
fdpost.combined <- subset(fdpost.combined, subset = predicted.celltype.l2 != "NK" &
                   predicted.celltype.l2 != "Eryth")
ncol(fdpost.combined)
```


# 7) Filter Additional RBC's  

## VlnPlots of HBA1, HBA2, & HBB
(linear scale)
```{r}
VlnPlot(fdpost.combined, features = c("HBA1", "HBA2", "HBB"), slot = "counts") + NoLegend()
```
(log scale)
```{r}
VlnPlot(fdpost.combined, features = c("HBA1", "HBA2", "HBB"), slot = "counts", log = T) + NoLegend()
```

Filter cells w/ less than or equal to 3 HBA1, HBA2, or HBB transcripts.  
```{r}
# Total number of cells in the object
ncol(fdpost.combined)

DefaultAssay(fdpost.combined) <- "RNA"
fdpost.combined <- subset(fdpost.combined, subset = HBA1 <= 3, slot = 'counts')
fdpost.combined <- subset(fdpost.combined, subset = HBA2 <= 3, slot = 'counts')
fdpost.combined <- subset(fdpost.combined, subset = HBB <= 3, slot = 'counts')
DefaultAssay(fdpost.combined) <- "SCT"
fdpost.combined <- subset(fdpost.combined, subset = HBA1 <= 3, slot = 'counts')
fdpost.combined <- subset(fdpost.combined, subset = HBA2 <= 3, slot = 'counts')
fdpost.combined <- subset(fdpost.combined, subset = HBB <= 3, slot = 'counts')

# Total number of cells in the object
ncol(fdpost.combined)
```

# Cell Counts after Filtering  
```{r}
fdpost.combined
```

## Cell Total
```{r}
# Total number of cells in the object
ncol(fdpost.combined)
```

## Cell Counts per Sample  
```{r}
table(fdpost.combined$orig.ident)
```
## Cell Counts from hpca.main  
```{r}
table(fdpost.combined$hpca.main)
```
## Cell Counts from encode.main  
```{r}
table(fdpost.combined$encode.main)
```

## Save Object
```{r}
save(fdpost.combined, file='./Objects/fdpostcombined.rda')
```



